---
title: |
  | \vspace{5cm} \LARGE COVID-19 Data Analysis with R
author: "Sandro Mikautadze"
date: "Last compiled on `r format(Sys.time(), '%d/%m/%Y')`"
output:
  pdf_document:
    number_sections: true
---

\newpage
\tableofcontents
\newpage


```{r setup, include = FALSE}
library(tibble)
library(dplyr)
covid19_raw <- read.csv("raw_data/covid19.csv")
covid19_allstates <- covid19_raw %>% filter(Province_State == "All States")

covid19_allstates_daily <- covid19_allstates %>% select(Date,
                                                        Country_Region,
                                                        active,
                                                        hospitalizedCurr,
                                                        daily_tested,
                                                        daily_positive
                                                        )
covid19_allstates_daily_sum <- covid19_allstates_daily %>% 
                               group_by(Country_Region) %>%
                               summarise(tested = daily_tested %>% sum,
                                         positive = daily_positive %>% sum,
                                         active = active %>% sum,
                                         hospitalized = hospitalizedCurr %>% sum
                               ) %>%
                               arrange(-tested)
covid19_top10 <- head(covid19_allstates_daily_sum, 10)
countries <- covid19_top10 %>% pull(Country_Region)
tested_cases <- covid19_top10 %>% pull(tested)
positive_cases <- covid19_top10 %>% pull(positive)
names(tested_cases) <- countries
names(positive_cases) <- countries
positive_tested_ratio <- sort(positive_cases / tested_cases)
ratio_top4 <- positive_tested_ratio[length(positive_tested_ratio):
                                    (length(positive_tested_ratio) - 3)]
uk_data <- c(0.11326062 * 100, 1473672, 166909)
us_data <- c( 0.10861819 * 100, 17282363, 1877179)
tur_data <- c(0.08071172 * 100, 2031192, 163941)
it_data <- c(0.06152337 * 100, 4091291, 251710)
top4_matrix <- rbind(uk_data, us_data, tur_data, it_data)
rownames(top4_matrix) <- c("UK", "US", "Turkey", "Italy")
colnames(top4_matrix) <- c("Positive/Tested %", "Tested", "Positive")
```

\newpage

# Introduction

This is exactly the TLDR you were looking for! 

Before starting to read, I would like to specify that the type of analysis is purely **descriptive** and does not aim at understanding the correlations and causation of the results.

All code chunks in this report use variables defined in the scripts.

**Enjoy!**

## What is the project about?

The following project is an individual assignment taken from a course on [Dataquest](https://www.dataquest.io/).
The aim is to analyse a dataset collected from January 20th, 2020 to June 1st, 2020 taken from [Kaggle](https://www.kaggle.com/datasets/lin0li/covid19testing).

The research question is the following: **which countries have had the highest number of positive cases against the number of tests?**

# Data

## Raw data Overview

We first analyse the dimensions, the column names and the information provided by each column in the dataset: 

```{r}
glimpse(covid19_raw)
```

Looking at the dataset repository from Kaggle we have further information on the meaning of each column:

* `Date`: date of the data collection

* `Country_Region`: country names

* `Province_State`: state/province names; value "All States" is put when state/provincial level data is NA

* `positive`: cumulative number of positive cases reported

* `active`: number of actively cases on that day

* `hospitalized`: cumulative number of hospitalized cases reported

* `hospitalizedCurr`: number of actively hospitalized cases on that day

* `recovered`: cumulative number of recovered cases reported

* `death`: cumulative number of deaths reported

* `total tested`: cumulative number of tests conducted

**Note**: Not all columns in our version of the data frame were present in the data description on Kaggle at the time of download. This indicates that the downloaded file was not updated to a later version of the data set.
Yet, Dataquest gives the information provided by the other columns:

* `Continent_Name`: continent name

* `Two_Letter_Country_Code`: country codes

* `Country_Region`: country names

* `daily_tested`: number of tests conducted on the day; if daily data is unavailable, `daily_tested` is averaged across number of days in between

* `daily_positive`: number of positive cases reported on the day; if daily data is unavailable, `daily_positive` is averaged across number of days in between

## Raw Data Cleaning

The first clean that we do is remove the `Province_State` column. Indeed, it might create some unwanted bias, as it also gives information about the specific province. So, to make the data "nationalized", we avoid looking at precise regions, and consider only those rows with "All States" value:

```{r}
glimpse(covid19_allstates)
```

As we can see we're left with just 30% of the data we previously had.

The second thing that we need to be aware of is the "unit of measurement" adopted by each column. For example, there are factors with cumulative information, others with daily information. To better address our research question, we only consider the columns providing daily data, that is `Date`, `Country_Region`, `active`, `hospitalizedCurr`, `daily_tested`, `daily_positive`. We get the following refactored data frame:
```{r}
glimpse(covid19_allstates_daily)
```

Now our data is clean.

# Analysis

Based on our cleaned dataset, we notice that our data is collected on a daily basis. Therefore, we can find the ratio of the overall number of positive cases over the total number of tests performed in each country each day.
What We want to do is create a data set that groups `Country_Region` and aggregates data on all tests made, positive, active and hospitalized cases in the period of time that the dataset covers (i.e. from January 20th to June 1st).

Displaying the countries based on those that have done more tests, we get:

```{r echo = FALSE}
head(covid19_allstates_daily_sum, 15)
```

Now, the result can be easily found by dividing the `positive` cases over the `tested` column and ranking the result for each country. We are going to do it using vectors, extracted from the data frame, as an exercise.

We first extract a vector of the 10 countries that have done more tests, over which we are going to conclude the rest of the analysis:
```{r}
countries
```

**Note**: if we extract the vector on the whole cleaned data set we get that the length of the vector `countries` is 108, which means that we cover more than half of the world with our data. Yet we decide to restrict the study case to just the top 10 because we have more data on the tests made, and hence a better chance of having the correct data (i.e. a more accurate estimate). As a matter of fact, there might be countries with higher ratios, but more non sufficient data, making our data exploration biased (or at least not correct enough).

Then, we extract our two final vectors to run the analysis and assign to each key the name of the country. Thus, we get:
```{r}
tested_cases
```
```{r}
positive_cases
```


The conclusion lies in a simple division: `positive_cases / tested_cases`, which yields the following result:
```{r}
positive_tested_ratio
```

# Conclusion

Referring the research question, since there is no precise meaning in the word *highest*, we decide to choose the top 4 countries with the highest positive/tests ratio. By doing so, we get:
```{r echo = FALSE}
ratio_top4
```

Therefore the **top 4 countries** were (in order):

1. **United Kingdom** with a ratio of `0.11326062`

2. **United States** with a ratio of `0.10861819`

3. **Turkey** with a ratio of `0.08071172`

4. **Italy** with a ratio of `0.06152337 `

Their corresponding data is displayed in the next matrix:
```{r echo = FALSE}
top4_matrix
```

# Methods of Improvement and Comments

Even though the structure of this analysis was very clear and defined, the results obtained might not properly answer the question. As a matter of fact, the whole analysis was mainly focused on the countries that had more data available regarding the number of tests made. It follows that, on the one hand, we ended up having a pretty good estimate of the ratio for those countries with many swabs, on the other hand, we might have excluded other "candidate" countries, simply because they did not have a sufficient number of swabs made in this dataset.
From a purely statistical point of view, I cannot assess whether the obtained result has a true value or not, as I still don't have sufficient knowledge on this matter.
To improve the analysis, we could have tried to find a better assessment method for the countries to include in the analysis. 
Overall, I am satisfied with the analysis, being this my first project.
